The Fates of Merging Supermassive Black Holes and a Proposal for a New Class of X-Ray Sources

### 2020-08-18 paper.tex
A few minor style revisions.

\begin{abstract}
We perform N-body simulations on some of the most massive galaxies extracted from a cosmological simulation of hierarchical structure formation with total masses in the range $10^{12} M_{\sun} < M_{tot} < 3\times 10^{13} M_{\sun}$ from $4\geq z \geq 0$.  After galactic mergers, we track the dynamical evolution of the infalling black holes (BHs) around their host's central BHs.  From 11 different simulations, we find that, of the 86 infalling BHs with masses > $10^4 M_{\sun}$, 36 merge with their host's central BH, 13 are ejected from their host galaxy, and 37 are still orbiting at $z=0$.  Across all galaxies, 33 BHs are kicked to a higher orbit after close interactions with the central BH binary or multiple, after which only one of them merged with their hosts.  These orbiting BHs should be detectable by their anomalous (not Low Mass X-ray Binary) spectra.  The X-ray luminosities of the orbiting massive BHs at z=0 are in the range $10^{28}-10^{43}$ $\mathrm{erg}~\mathrm{s}^{-1}$, with a currently undetectable median value of $10^{33}$ $\mathrm{erg}~\mathrm{s}^{-1}$.  However, the most luminous $\sim$5\% should be detectable by existing X-ray facilities.
\end{abstract}

@@

Orbiting black holes have been studied in the past using both semi-analytical methods and dynamical simulations.  \citet{2003ApJ...582..559V} followed the merger history of dark matter haloes and associated BH mergers using Monte-Carlo realizations, and found a population of BHs orbiting in galaxy haloes and the intergalactic medium at the present epoch.  In a follow-up study, \citet{2005MNRAS.358..913V} found that the orbiting black holes should have luminosities in X-ray $\lessapprox10^9 M_{\sun}$.  \citet{2008MNRAS.390.1311B} looked at the effects of gravitational recoil due to BH mergers on the growth of SMBHs, and found that the recoiled BHs can orbit their host galaxy for $10^6-10^9$ years before returning to the center, while accreting gas equivalent to approximate $10{\%}$ of their initial mass.  \citet{2017MNRAS.472.1526C} simulated recoiling black hole trajectories, altering various parameters in their simulations and observing the effects on escape velocities.  They found that the overall mass of the host halo and the accretion rate onto the halo are the largest factors affecting the escape velocity of a recoiled BH.  In a large-scale cosomological simulation (ROMULUS25) of Milky Way-mass galaxies, \citet{2018ApJ...857L..22T} found that such haloes would have approximately 5 SMBHs within $10kpc$ from the galactic center, and 12 SMBHs within they virial radius.  \citet{2020MNRAS.495.4681I} used a semi-analytical model to study the mass assembly and spin evolution of SMBHs, as well as the effects of gravitational recoil on central and orbiting black holes.  One of their findings is that the number and masses of orbiting black holes corresponds with the halo and stellar mass of the host galaxy, with their location within the galaxy being approximately $0.5R_{200}$.

@@

The merger tree data from \citet{2012MNRAS.425..641L} and \citet{2015ApJ...799..178K} provided 1,830 galaxies in total.  However, not all the galaxies were suitable for our simulations.  We placed further requirements as follows:
\begin{itemize}
\item The galaxies had to exist through the entire simulation ($4\geq z \geq 0$).  If they merged with other galaxies, they had to have been the "surviving" galaxy at each merger.
\item They had to have accumulated orbiting black holes by $z = 0$.
\end{itemize}

The vast majority of the full set of galaxies (1,537) were eliminated because they did not come into existence until after $z=4$.  Of the remaining 293 galaxies, 67 of them existed at z=0, and 51 of them also had black holes orbiting at z=0.  In considering time and resources, we chose the 11 galaxies that had the largest number of black holes to run the simulations.  The black holes in these galaxies accounted for nearly $84{\%}$ of all orbiting black holes across all the galaxies, so it represented a reasonable balance between completeness and efficiency.

@@

\begin{table}
\begin{center}
\caption{Galaxy characteristics at z=0. $M_\mathrm{DM}$ is dark matter mass.  $M_\mathrm{CBH}$ is central black hole mass.  Units are $M_{\odot}$.}
\begin{tabular} {|c | c | c| c|}
\hline
$M_\mathrm{DM}$ & $M_\mathrm{CBH}$ & $M_{*} \mathrm{(orig.)}$ & $M_{*} \mathrm{(rescaled)}$ \\
\hline
 2.85 $\times 10^{13}$ 	&	 1.46 $\times 10^9$	&	5.58 $\times 10^{12}$ 	&	3.30 $\times 10^{11}$  \\
 2.48 $\times 10^{13}$ 	&	 2.40 $\times 10^9$	&	4.80 $\times 10^{12}$ 	&	2.25 $\times 10^{11}$  \\
 2.29 $\times 10^{13}$ 	&	 1.22 $\times 10^9$	&	4.46 $\times 10^{12}$ 	&	2.68 $\times 10^{11}$  \\
 1.18 $\times 10^{13}$ 	&	 3.62 $\times 10^9$	&	2.32 $\times 10^{12}$ 	&	2.83 $\times 10^{11}$  \\
 1.03 $\times 10^{13}$ 	&	 9.85 $\times 10^9$	&	1.96 $\times 10^{12}$ 	&	1.62 $\times 10^{11}$  \\
 1.00 $\times 10^{13}$ 	&	 9.32 $\times 10^9$	&	1.95 $\times 10^{12}$ 	&	2.49 $\times 10^{11}$  \\
 8.92 $\times 10^{12}$ 	&	 5.60 $\times 10^9$	&	1.74 $\times 10^{12}$ 	&	3.20 $\times 10^{11}$  \\
 8.56 $\times 10^{12}$ 	&	 6.86 $\times 10^9$	&	1.62 $\times 10^{12}$ 	&	1.83 $\times 10^{11}$  \\
 6.94 $\times 10^{12}$ 	&	 2.02 $\times 10^9$	&	1.72 $\times 10^{12}$ 	&	1.51 $\times 10^{11}$  \\
 6.29 $\times 10^{12}$ 	&	 1.22 $\times 10^9$	&	1.23 $\times 10^{12}$ 	&	1.87 $\times 10^{11}$  \\
 1.84 $\times 10^{12}$ 	&	 2.21 $\times 10^9$	&	5.52 $\times 10^{11}$ 	&	6.42 $\times 10^{10}$  \\
\hline
\end{tabular}
\end{center}
\label{table:gal_char}
\end{table}

@@

\begin{figure}
\begin{center}
\includegraphics[width=1.0\columnwidth]{plots/Summary_Results_All_Galaxies.png}
\caption{Histogram of merged, ejected, and orbiting SMBHs at z=0 as a function of BH mass.  Ejected black holes span almost the entire mass range, while lower-mass black holes are orbiting and higher mass black holes tended to merge.  Colored arrows along upper x-axis represent median masses of respective sets.  Histogram bars are stacked.}
\label{fig:meosmbh}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=1.0\columnwidth]{plots/kicked_stats.png}
\caption{Histogram of kicked and not-kicked SMBHs at z=0.  Colored arrows along upper x-axis represent median masses of respective sets.  Histograms are stacked.}
\label{fig:kicked_stats}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=1.0\columnwidth]{plots/Final_Radius_v_Mass_of_Orbiting_BHs.png}
\caption{Final radii of remaining orbital black holes at z=0.  Kicked black holes come back to the center, but due to their highly eccentric nature, they spend most of their time far from the center.  There is a slight inverse relationship between BH mass and final distance from center (solid line).}
\label{fig:mvr}
\end{center}
\end{figure}

@@

Figure \ref{fig:ldobhs} represents a histogram of the final X-ray luminosities of the black holes, overlaid by a lognormal distribution, obtained with a mean $\mu=1.0\times10^{33}\,\text{erg s}^{-1}$ and standard deviation $\sigma=4.5\times10^{3}\,\text{erg s}^{-1}$.
\begin{figure}
\begin{center}
\includegraphics[width=1.0\columnwidth]{plots/Final_Luminosities_using_Kohei_factor.png}
\caption{X-ray thermal luminosity distribution of orbiting BHs at z=0.  Green region under curve represents approximate range of detectable orbiting black holes, with the lower limit being approximately $10^{38}\;\text{erg}\;\text{s}^{-1}$ (\citet{2018ApJ...862...73L}). The dashed, red vertical line is the median X-ray thermal luminosity of our sources ($1.0\times10^{33}\;\text{erg}\;\text{s}^{-1}$).}
\label{fig:ldobhs}
\end{center}
\end{figure}

@@

\section{Conclusions}\label{sec:conclusions}
We performed N-body simulations on 11 galaxies ranging in mass between $10^{12}$ M$_{\sun}$ < M < 3$\times$ $10^{13}$ M$_{\sun}$, from 4 $\geq$ z $\geq$ 0.  A total of 86 orbiting black holes were followed in our simulations, resulting in 42$\%$ mergers with their host galaxy's resident black hole, while another 43$\%$ remained in orbit at $z=0$.  Those remaining orbiting black holes had X-ray luminosities in the range of $10^{28}-10^{43}\mathrm{erg}~\mathrm{s}^{-1}$.  The median luminosity of $10^{33}\mathrm{erg}~\mathrm{s}^{-1}$ would be undetectable with present instruments, although the top $5\%$ most luminous from this set could be detectable.  If detected, these orbiting massive black holes would constitute a new category of X-ray source.

>>>>

\begin{abstract}
We perform N-body simulations on some of the most massive galaxies extracted from a cosmological simulation of hierarchical structure formation with total masses in the range $10^{12} M_{\sun} < M_{tot} < 3\times 10^{13} M_{\sun}$ from $4\geq z \geq 0$.  After galactic mergers, we track the dynamical evolution of the infalling black holes (BHs) around their host's central BHs.  From 11 different simulations, we find that, of the 86 infalling BHs with masses > $10^4 M_{\sun}$, 36 merge with their host's central BH, 13 are ejected from their host galaxy, and 37 are still orbiting at $z=0$.  Across all galaxies, 33 BHs are kicked to a higher orbit after close interactions with the central BH binary or multiple, after which only one of them merged with their hosts.  These orbiting BHs should be detectable by their anomalous (not Low Mass X-ray Binary) spectra.  The X-ray luminosities of the orbiting massive BHs at $z=0$ are in the range $10^{28}-10^{43}$ $\mathrm{erg}~\mathrm{s}^{-1}$, with a currently undetectable median value of $10^{33}$ $\mathrm{erg}~\mathrm{s}^{-1}$.  However, the most luminous $\sim$5\% should be detectable by existing X-ray facilities.
\end{abstract}

@@

Orbiting black holes have been studied in the past using both semi-analytical methods and dynamical simulations.  \citet{2003ApJ...582..559V} followed the merger history of dark matter haloes and associated BH mergers using Monte-Carlo realizations, and found a population of BHs orbiting in galaxy haloes and the intergalactic medium at the present epoch.  In a follow-up study, \citet{2005MNRAS.358..913V} found that the orbiting black holes should have luminosities in X-ray $\lessapprox10^{39} M_{\sun}$.  \citet{2008MNRAS.390.1311B} looked at the effects of gravitational recoil due to BH mergers on the growth of SMBHs, and found that the recoiled BHs can orbit their host galaxy for $10^6-10^9$ years before returning to the center, while accreting gas equivalent to approximate $10{\%}$ of their initial mass.  \citet{2017MNRAS.472.1526C} simulated recoiling black hole trajectories, altering various parameters in their simulations and observing the effects on escape velocities.  They found that the overall mass of the host halo and the accretion rate onto the halo are the largest factors affecting the escape velocity of a recoiled BH.  In a large-scale cosomological simulation (ROMULUS25) of Milky Way-mass galaxies, \citet{2018ApJ...857L..22T} found that such haloes would have approximately 5 SMBHs within $10kpc$ from the galactic center, and 12 SMBHs within they virial radius.  \citet{2020MNRAS.495.4681I} used a semi-analytical model to study the mass assembly and spin evolution of SMBHs, as well as the effects of gravitational recoil on central and orbiting black holes.  One of their findings is that the number and masses of orbiting black holes corresponds with the halo and stellar mass of the host galaxy, with their location within the galaxy being approximately $0.5R_{200}$.

@@

The merger tree data from \citet{2012MNRAS.425..641L} and \citet{2015ApJ...799..178K} provided 1,830 galaxies in total.  However, not all the galaxies were suitable for our simulations.  We placed further requirements as follows:
\begin{itemize}
\item The galaxies had to exist through the entire simulation ($4\geq\;z\;\geq 0$).  If they merged with other galaxies, they had to have been the "surviving" galaxy at each merger.
\item They had to have accumulated orbiting black holes by $z = 0$.
\end{itemize}

The vast majority of the full set of galaxies (1,537) were eliminated because they did not come into existence until after $z=4$.  Of the remaining 293 galaxies, 67 of them existed at $z=0$, and 51 of them also had black holes orbiting at $z=0$.  In considering time and resources, we chose the 11 galaxies that had the largest number of black holes to run the simulations.  The black holes in these galaxies accounted for nearly $84{\%}$ of all orbiting black holes across all the galaxies, so it represented a reasonable balance between completeness and efficiency.

@@

\begin{table}
\begin{center}
\caption{Galaxy characteristics at $z=0$. $M_\mathrm{DM}$ is dark matter mass.  $M_\mathrm{CBH}$ is central black hole mass.  Units are $M_{\odot}$.}
\begin{tabular} {|c | c | c| c|}
\hline
$M_\mathrm{DM}$ & $M_\mathrm{CBH}$ & $M_{*} \mathrm{(orig.)}$ & $M_{*} \mathrm{(rescaled)}$ \\
\hline
 2.85 $\times 10^{13}$ 	&	 1.46 $\times 10^9$	&	5.58 $\times 10^{12}$ 	&	3.30 $\times 10^{11}$  \\
 2.48 $\times 10^{13}$ 	&	 2.40 $\times 10^9$	&	4.80 $\times 10^{12}$ 	&	2.25 $\times 10^{11}$  \\
 2.29 $\times 10^{13}$ 	&	 1.22 $\times 10^9$	&	4.46 $\times 10^{12}$ 	&	2.68 $\times 10^{11}$  \\
 1.18 $\times 10^{13}$ 	&	 3.62 $\times 10^9$	&	2.32 $\times 10^{12}$ 	&	2.83 $\times 10^{11}$  \\
 1.03 $\times 10^{13}$ 	&	 9.85 $\times 10^9$	&	1.96 $\times 10^{12}$ 	&	1.62 $\times 10^{11}$  \\
 1.00 $\times 10^{13}$ 	&	 9.32 $\times 10^9$	&	1.95 $\times 10^{12}$ 	&	2.49 $\times 10^{11}$  \\
 8.92 $\times 10^{12}$ 	&	 5.60 $\times 10^9$	&	1.74 $\times 10^{12}$ 	&	3.20 $\times 10^{11}$  \\
 8.56 $\times 10^{12}$ 	&	 6.86 $\times 10^9$	&	1.62 $\times 10^{12}$ 	&	1.83 $\times 10^{11}$  \\
 6.94 $\times 10^{12}$ 	&	 2.02 $\times 10^9$	&	1.72 $\times 10^{12}$ 	&	1.51 $\times 10^{11}$  \\
 6.29 $\times 10^{12}$ 	&	 1.22 $\times 10^9$	&	1.23 $\times 10^{12}$ 	&	1.87 $\times 10^{11}$  \\
 1.84 $\times 10^{12}$ 	&	 2.21 $\times 10^9$	&	5.52 $\times 10^{11}$ 	&	6.42 $\times 10^{10}$  \\
\hline
\end{tabular}
\end{center}
\label{table:gal_char}
\end{table}

@@

\begin{figure}
\begin{center}
\includegraphics[width=1.0\columnwidth]{plots/Summary_Results_All_Galaxies.png}
\caption{Histogram of merged, ejected, and orbiting SMBHs at $z=0$ as a function of BH mass.  Ejected black holes span almost the entire mass range, while lower-mass black holes are orbiting and higher mass black holes tended to merge.  Colored arrows along upper x-axis represent median masses of respective sets.  Histogram bars are stacked.}
\label{fig:meosmbh}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=1.0\columnwidth]{plots/kicked_stats.png}
\caption{Histogram of kicked and not-kicked SMBHs at $z=0$.  Colored arrows along upper x-axis represent median masses of respective sets.  Histograms are stacked.}
\label{fig:kicked_stats}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=1.0\columnwidth]{plots/Final_Radius_v_Mass_of_Orbiting_BHs.png}
\caption{Final radii of remaining orbital black holes at $z=0$.  Kicked black holes come back to the center, but due to their highly eccentric nature, they spend most of their time far from the center.  There is a slight inverse relationship between BH mass and final distance from center (solid line).}
\label{fig:mvr}
\end{center}
\end{figure}

@@

Figure \ref{fig:ldobhs} represents a histogram of the final X-ray luminosities of the black holes, overlaid by a lognormal distribution, obtained with a mean $\mu=1.0\times10^{33}\,\text{erg s}^{-1}$ and standard deviation $\sigma=4.5\times10^{3}\,\text{erg s}^{-1}$.
\begin{figure}
\begin{center}
\includegraphics[width=1.0\columnwidth]{plots/Final_Luminosities_using_Kohei_factor.png}
\caption{X-ray thermal luminosity distribution of orbiting BHs at $z=0$.  Green region under curve represents approximate range of detectable orbiting black holes, with the lower limit being approximately $10^{38}\;\text{erg}\;\text{s}^{-1}$ (\citet{2018ApJ...862...73L}). The dashed, red vertical line is the median X-ray thermal luminosity of our sources ($1.0\times10^{33}\;\text{erg}\;\text{s}^{-1}$).}
\label{fig:ldobhs}
\end{center}
\end{figure}

@@

\section{Conclusions}\label{sec:conclusions}
We performed N-body simulations on 11 galaxies ranging in mass between $10^{12}$ M$_{\sun}$\;<\;M\;<\;3$\times$ $10^{13}$ M$_{\sun}$, from $4\;\geq\;z\;\geq\;0$.  A total of 86 orbiting black holes were followed in our simulations, resulting in 42$\%$ mergers with their host galaxy's resident black hole, while another 43$\%$ remained in orbit at $z=0$.  Those remaining orbiting black holes had X-ray luminosities in the range of $10^{28}-10^{43}\mathrm{erg}~\mathrm{s}^{-1}$.  The median luminosity of $10^{33}\mathrm{erg}~\mathrm{s}^{-1}$ would be undetectable with present instruments, although the top $5\%$ most luminous from this set could be detectable.  If detected, these orbiting massive black holes would constitute a new category of X-ray source.
